Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-02-02 reports/manuscript/paper.tex
Adjust figure placement

\begin{abstract}
    % Abstracts should be unreferenced and typically around 150 words
    The white matter contains long-range connections between different
    brain regions and the organization of these connections holds important
    implications for brain function in health and disease. Tractometry
    uses diffusion-weighted magnetic resonance imaging (dMRI)
    to quantify tissue properties (e.g. fractional anisotropy (FA),
    mean diffusivity (MD), etc.), along the trajectories of these
    connections.
    % \cite{yeatman2012tract}.
    Statistical inference from
    tractometry usually either averages these quantities along
    the length of each bundle in each individual or performs analysis
    point-by-point along each bundle, with group comparisons or regression
    models computed separately for each point along every one of the bundles.
    These approaches are limited in their sensitivity, in the former case, or
    in their statistical power, in the latter.
    In the present work, we developed a method based on the sparse group
    lasso (SGL)
    % \cite{simon2013sparse}
    that takes into account tissue
    properties measured along all of the bundles, and selects informative
    features by enforcing sparsity, not only at the level of individual
    bundles, but also across the entire set of bundles and all of the measured
    tissue properties.
    % The sparsity penalties for each of these constraints
    % is identified using a nested cross-validation scheme that guards
    % against over-fitting and simultaneously identifies the correct
    % level of sparsity.
    We demonstrate the accuracy of the method in two
    settings: i) In a classification setting, patients with amyotrophic
    lateral sclerosis (ALS) are accurately distinguished from matched
    controls.
    % \cite{sarica2017corticospinal}.
    Furthermore, SGL automatically
    identifies FA in the corticospinal tract as important for this
    classification, correctly finding the parts of the white matter known
    to be affected by the disease. ii) In a regression setting, dMRI is
    used to accurately predict ``brain age.''
    % \cite{yeatman2014lifespan, Brown2012-so}.
    In this case, the weights are distributed throughout the
    white matter indicating that many different regions of the white matter
    change with development and contribute to the prediction of age. Thus,
    SGL makes it possible to leverage the multivariate relationship between
    diffusion properties measured along multiple bundles to make accurate
    predictions of subject characteristics while simultaneously discovering
    the most relevant features of the white matter for the characteristic of
    interest.
\end{abstract}

@@

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 978, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the tractometry-extracted
features FA and MD in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} for the HBN and Cam-CAN datasets.

@@

\printfigures

>>>>

\begin{abstract}
    The white matter contains long-range connections between different brain
    regions and the organization of these connections holds important
    implications for brain function in health and disease. Tractometry uses
    diffusion-weighted magnetic resonance imaging (dMRI) to quantify tissue
    properties along the trajectories of these connections.
    % \cite{yeatman2012tract}.
    Statistical inference from tractometry usually either averages these
    quantities along the length of each fiber bundle or computes regression
    models separately for each point along every one of the bundles. These
    approaches are limited in their sensitivity, in the former case, or in
    their statistical power, in the latter. We developed a method based on
    the sparse group lasso (SGL)
    % \cite{simon2013sparse}
    that takes into account tissue properties along all of the bundles and
    selects informative features by enforcing both global and bundle-level
    sparsity. We demonstrate the performance of the method in two settings:
    i) in a classification setting, patients with amyotrophic lateral
    sclerosis (ALS) are accurately distinguished from matched controls.
    % \cite{sarica2017corticospinal}.
    Furthermore, SGL identifies the corticospinal tract as important for this
    classification, correctly finding the parts of the white matter known to
    be affected by the disease. ii) In a regression setting, SGL accurately
    predicts ``brain age.''
    % \cite{yeatman2014lifespan, Brown2012-so}.
    In this case, the weights are distributed throughout the white matter
    indicating that many different regions of the white matter change over
    the lifespan. Thus, SGL leverages the multivariate relationships between
    diffusion properties in multiple bundles to make accurate phenotypic
    predictions while simultaneously discovering the most relevant features
    of the white matter.
\end{abstract}

@@

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 978, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the tractometry-extracted
features FA and mean diffusivity (MD) in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} for the HBN and Cam-CAN datasets.

@@

### 2021-02-02 reports/manuscript/supplemental.tex
Adjust figure placement

@@

\begin{figure}[!h]
    \includegraphics[width=\columnwidth]{nested-cross-validation.pdf}
    \caption{{\bf Nested cross-validation}
        \label{fig:nested-cross-val}
        We evaluate model quality using a nested $k$-fold cross validation
        scheme. At level-0, the input data is decomposed into $k_0$ shuffled
        groups and optimal hyperparameters are found for the level-0 training
        set. To avoid overfitting, the optimal hyperparameters are themselves
        evaluated using a cross-validation scheme taking place at level-1 of
        the decomposition, where each level-0 training set is further
        decomposed into $k_1 = 3$ shuffled groups. In the classification
        case, the training and test splits are stratified by diagnosis. For
        the ALS and WH data, $k_0 = 10$, while for the HBN and Cam-CAN data,
        $k_0 = 5$.
    }
\end{figure}

@@

To objectively evaluate model performance and guard against over-fitting,
we used a nested cross-validation scheme, which is depicted for the binary
classification case in \cref{fig:nested-cross-val}. The subjects (i.e. rows
of the feature matrix $\mathbf{X}$ in \cref{main:fig:methods:group-structure}
and \cref{eq:lm}) are first shuffled and then decomposed into $k_0$ batches,
hereafter referred to as folds. For the ALS and WH datasets, we used $k_0 =
10$ and for the HBN and Cam-CAN datasets, $k_0 = 5$. For each unique fold, we
hold that fold out as the test\textsubscript{0} set and let the remaining
data comprise the train\textsubscript{0} set, with the subscript indicating
the depth of the nested decomposition. We further decompose each
train\textsubscript{0} set into three folds, and again for each unique fold,
we hold out that fold as the test\textsubscript{1} set and let the remaining
data comprise the train\textsubscript{1} set. At level-1 of the
decomposition, we fit an SGL model using fixed regularization meta-parameters
$\alpha$ and $\lambda$, training the model using train\textsubscript{1} and
evaluating the fit on test\textsubscript{1}. We find the optimal values for
$\alpha$ and $\lambda$ using sequential model-based optimization, implemented
using the scikit-optimize \verb|BayesSearchCV| class\cite{scikit-optimize}.
For continuous numerical $y$, \verb|BayesSearchCV| searches for
meta-parameter values that maximize the $R^2$ averaged over test sets. For
binary categorical $y$ \verb|BayesSearchCV| seeks to maximize the
classification accuracy. Using hyperoptimization, we find optimal
regularization parameters and $\hat{\beta}$ for each train\textsubscript{0}
set and then use those to predict values for data in test\textsubscript{0}.
Thus each subject in the dataset has a predicted phenotype derived from a
model that never saw its particular subject's data. In the classification
case, the shuffling into folds is stratified such that each fold has a
population that preserves the percentage of each class found in the larger
input data.

@@

The full software implementation of the SGL approach presented here is available
in a Python software package called \texttt{AFQ-Insight}, which is developed
publicly in \url{https://github.com/richford/afq-insight}.
The version of the code used to produce the results herein is also available in
\url{https://doi.org/10.5281/zenodo.3585942}\todo{Update version here}.
AFQ-Insight reads the target and feature data that has been processed by AFQ
from comma separated value (CSV) files conforming to the AFQ-Browser data
format \cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python library
\cite{mckinney2010data}. The software provides different options for imputing
missing data in the feature matrix. Missing interior nodes are imputed using
linear interpolation. For missing exterior nodes, the user may choose between
linear extrapolation and constant forward- or back-fill. Imputation uses only
values from adjacent nodes in the same white matter bundle in the same
subject so there is no danger of data leakage from other subjects. It uses
the scikit-learn \cite{scikit-learn} library to decompose input data into
separate test and train datasets, to scale each feature to have zero mean and
unit variance, and to evaluate each fit in the hyperparameter search using
appropriate classification and regression metrics such as accuracy, area
under the receiver operating curve (AUC ROC), and coefficient of
determination ($R^2$). Internally, \texttt{AFQ-Insight} uses the \texttt{groupyr} software library \cite{groupyr} mentioned above to solve the SGL.

>>>>

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{nested-cross-validation.pdf}
    \caption{{\bf Nested cross-validation}
        \label{fig:nested-cross-val}
        We evaluate model quality using a nested $k$-fold cross validation
        scheme. At level-0, the input data is decomposed into $k_0$ shuffled
        groups and optimal hyperparameters are found for the level-0 training
        set. To avoid overfitting, the optimal hyperparameters are themselves
        evaluated using a cross-validation scheme taking place at level-1 of
        the decomposition, where each level-0 training set is further
        decomposed into $k_1 = 3$ shuffled groups. In the classification
        case, the training and test splits are stratified by diagnosis. For
        the ALS and WH data, $k_0 = 10$, while for the HBN and Cam-CAN data,
        $k_0 = 5$.
    }
\end{figure}

@@

@@

To objectively evaluate model performance and guard against over-fitting,
we used a nested cross-validation scheme, which is depicted for the binary
classification case in \cref{fig:nested-cross-val}. The subjects (i.e. rows
of the feature matrix $\mathbf{X}$ in \cref{main:fig:methods:group-structure}
and \cref{eq:lm}) are first shuffled and then decomposed into $k_0$ batches,
hereafter referred to as folds. For the ALS and WH datasets, we used $k_0 =
10$ and for the HBN and Cam-CAN datasets, $k_0 = 5$. For each unique fold, we
hold that fold out as the test\textsubscript{0} set and let the remaining
data comprise the train\textsubscript{0} set, with the subscript indicating
the depth of the nested decomposition. We further decompose each
train\textsubscript{0} set into three folds, and again for each unique fold,
we hold out that fold as the test\textsubscript{1} set and let the remaining
data comprise the train\textsubscript{1} set. At level-1 of the
decomposition, we fit an SGL model using fixed regularization meta-parameters
$\alpha$ and $\lambda$, training the model using train\textsubscript{1} and
evaluating the fit on test\textsubscript{1}. We find the optimal values for
$\alpha$ and $\lambda$ using sequential model-based optimization, implemented
using the scikit-optimize \verb|BayesSearchCV| class \cite{scikit_optimize}.
For continuous numerical $y$, \verb|BayesSearchCV| searches for
meta-parameter values that maximize the $R^2$ averaged over test sets. For
binary categorical $y$ \verb|BayesSearchCV| seeks to maximize the
classification accuracy. Using hyperoptimization, we find optimal
regularization parameters and $\hat{\beta}$ for each train\textsubscript{0}
set and then use those to predict values for data in test\textsubscript{0}.
Thus each subject in the dataset has a predicted phenotype derived from a
model that never saw its particular subject's data. In the classification
case, the shuffling into folds is stratified such that each fold has a
population that preserves the percentage of each class found in the larger
input data.

@@

The full software implementation of the SGL approach presented here is available
in a Python software package called \texttt{AFQ-Insight}, which is developed
publicly in \url{https://github.com/richford/afq-insight}.
The version of the code used to produce the results herein is also available in
\url{https://doi.org/10.5281/zenodo.4316000}.
AFQ-Insight reads the target and feature data that has been processed by AFQ
from comma separated value (CSV) files conforming to the AFQ-Browser data
format \cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python library
\cite{mckinney2010data}. The software provides different options for imputing
missing data in the feature matrix. Missing interior nodes are imputed using
linear interpolation. For missing exterior nodes, the user may choose between
linear extrapolation and constant forward- or back-fill. Imputation uses only
values from adjacent nodes in the same white matter bundle in the same
subject so there is no danger of data leakage from other subjects. It uses
the scikit-learn \cite{scikit-learn} library to decompose input data into
separate test and train datasets, to scale each feature to have zero mean and
unit variance, and to evaluate each fit in the hyperparameter search using
appropriate classification and regression metrics such as accuracy, area
under the receiver operating curve (AUC ROC), and coefficient of
determination ($R^2$). Internally, \texttt{AFQ-Insight} uses the \texttt{groupyr} software library \cite{groupyr} mentioned above to solve the SGL.
